\chapter{Thesis conclusion}
\label{chap:thesis-ccl}

%\addcontentsline{toc}{chapter}{Thesis conclusion}

Through this thesis, several points have been raised.
The focus on the two case studies are believed to be representative of the current privacy issues.

\section{Android ecosystem has to evolve}

The Android system is not technically unsecured with the exception of some unavoidable security flaws.
However the security issues are present and explained by two factors.\\

One of them comes from the negligence or ignorance of the security risks during the installation process.
Users will install malicious applications that could have been avoided.
Applications similar to DroidWatcher are technically possible but the risk of propagation could be easily minimised with a better security awareness.

However, the user is not the only one responsible for security issues.
In comparison with the model used in iOS, the open politic of Android is also responsible for the higher risk.
An open politic has indubitable advantages for the user's freedom but enables abuses and reduces the marge of action of Google (release of updates, use of Play Store, etc.) left to manufacturer responsibility.
The permission system also is not perfect, it lacks from precision, clarity and enables too easily abusive behaviours.\\

It is questionable if Android will be able to react to the recent strong raise in malicious applications.
Will antivirus software be mandatory as on the Windows platform or will a closed and certified marketplace such as the App Store be developed?
It is however imaginable that an evolution in both the user's behaviour and the platform security measures will happen to adapt to this risk.
The fragmentation of the market should also be resolved by a better reaction to security threads to manufacturer (better support of older models, etc.).

\section{Wireless camera must be shipped secured}

The analysis of the DCS-2130 reveals several security issues that can be exploited.
The use of such camera without taking additional security measures is considered as unsecured and unadvised.
It is possible to use the camera in a secure way but it requires a knowledge of security principles that most users do not have.\\

These security issues are not impossible to fix but it requires the manufacturer to dedicate more attention into security.
The security of appliances such as a wireless camera needs to be present \emph{out-of-the-box}, not requiring additional effort from the end-user.

\section{Privacy does matter}

On both Android systems and the wireless cameras, the potential leak of information has very dangerous consequences.
A smartphone is able to easily track every movement of a user or reveal sensible private information.
The case of camera abuses is easily imaginable and harmful as the burglar scenario showed.\\

The privacy is an important aspect of the use of technology.
If the state of care for privacy does not improve, scandals such as the ones related to the cache database in iPhone and Android systems or the TRENDnet hack will happen again.

\section{Beware of the convenience}

The common aspects of these technologies is the convenience that drives their adoption.
As these technologies are used as a service, they are only considered for the tasks they achieve and not as the systems they are.
Today's tendency is to abstract all the complexity from the user and sell a device that \emph{just works}.
However this is often untrue, a system is more complex than it is advertised.
Security and privacy are some of the limitations of this abstraction.

% Considering a device as a function brings risky behaviours in term of management of this device.
% It is very uncommon to apply specific security measures or to update such devices.
% Manufacturers do not necessarily provide updates of their devices as it is often seen in Android devices.
% If it is not securely configured by default, it will probably stay that way during all the device usage.

\section{Personal thoughts}

I found the researches and analysis made during this thesis very interesting.
The redaction of this document helped me to develop a critical view of the technologies we use every day.
The analysed case studies reveal various security issues and enable potential abuses in terms of privacy that I did not suspect.
I hope that reading this thesis will create a better awareness of the fact that these technologies are not perfect and that relying blindly on them is not without risk.
Fixing the security and privacy issues is not possible in one day, it starts with a better concern in these issues from users and manufacturers.
It is clear that this improvement is only possible with intensified regulation efforts on the way to handle the privacy of customers.